---
title: "WDFW Age Composition Bootstrap Estimator"
format: 
  html:
    toc: true
    theme: sketchty
    dark: darkly
    embed-resources: true
    self-contained-math: true
    code-tools: true
    code-fold: true
editor: visual
echo: true
params:
  current_year: 2024
  sp: "Chum" #"Chinook" #"Coho" "Chum" "Pink"
  system: "Wallace River" #"Puyallup River"
---

|                                                 |
|:------------------------------------------------|
| Last Updated `r format(Sys.time(), '%m/%d/%Y')` |

\
\

## Introduction

This script contains a bootstrap estimator for otolith samples collected where various hypothetical sampling rate scenarios (10-100% by 10% increments) are explored. Simulated age compositions are produced for a number of bootstrap samples designated by the user by stepping through the reported sampling dates for each year included and sampling otolith ages at the specified rate for each treatment. Effective sample size is then calculated for each sampling rate treatment and summarized as the harmonic mean (Stewart & Hamel, 2014; Eq. 6) across bootstrap replicates.

## Setup, pull data, and format

Use this section to load libraries, set directories, and designate any life history or geographic groupings. Then read in the data, calculate brood year returns, and format for model fitting.

```{r setup, message=FALSE, warning=FALSE}
#####################################################################################################################################################
### Clear workspace
# rm(list=ls())

### Load libraries
library(ggplot2);library(ggh4x);library(ggthemes);library(ggeffects);library(kableExtra);
library(tidyr);library(dplyr);library(data.table);library(purrr);library(readr)
library(lubridate);library(janitor);library(stringi);library(rmdformats);
library(nlme);library(MuMIn);library(Hmisc);library(mgcv);library(scales);
library(odbc);library(RODBC);library(DBI);library(stringr);library(mgcv);
library(RJSONIO);library(RCurl);library(httr);library(jsonlite);library(glmmTMB);

### Set up drives, seed, & controls
drive          <-paste0("C:/Users/sism1477/OneDrive - Washington State Executive Branch Agencies/Desktop/wdfw_age_be/")
`%nin%`        <-negate(`%in%`) # create an exclude operator
# source(paste(drive,"/Functions/Functions.R",sep="")) #Source seed function to create seedlist; only run once then turn off
load(paste(drive,"/seed.list_v2",sep="")) #load seed folder again
bootest.seed <- as.numeric(seed.list[[1]][,"bootest.seed3"]) #read recruitment seed from seedlist function
n_boot <-1000

### Species designations
sp_name_list  <-c("Chum") #, "chinook", "pink", "coho", "sockeye", "steelhead", "cutthroat", "rainbow", "bull trout"
# sp_code_list <-c() # "1", "2", "5", "6"

### Stage designations
# sy_stages    <-c("E","F")
# yrl_stages   <-c("G","Y")
# smolt_stages <-c("P","S")

### Read-in age data & calc real age comps
chum24                    <-read.csv(paste0(drive,"chum_ages_wallace_2024.csv"))
chum20_23                 <-read.csv(paste0(drive,"chum_ages_wallace_2020-2023.csv"))
age_data                  <-rbind(chum24[,c("species_description","location_name","sample_datetime", "gilbert_rich")],
                                  chum20_23[,c("species_description","location_name","sample_datetime", "gilbert_rich")])
age_data                  <-filter(age_data, gilbert_rich!=c("Unreadable"))
age_data$return_yr        <-as.character(substr(age_data$sample_datetime,1,4))
age_data$return_yr        <-as.numeric(age_data$return_yr)
age_data$month_date       <-as.character(substr(age_data$sample_datetime,6,10))
age_data$gilbert_rich_int <-substr(chartr("₀₁₂₃₄₅₆₇₈₉", "0123456789", age_data$gilbert_rich),1,1)
age_data$gilbert_rich_int <-as.numeric(age_data$gilbert_rich_int)

```

## Resample Age Data

```{r boot}
### Set up bootstrapping parameters
age_comps_real_sp <-list()
age_comps_real_sp_y <-list()
age_comps_sim_trt <-list()
age_comps_sim_sp <-list()
age_data_sim <-list()
n_nom_real_sp <-list()
n_nom_sim_sp <-list()
samp_rate <-c(seq(0.1,1,0.1))
# print(Sys.time())
# sp=b=s=1
# y=4

for(sp in 1:length(sp_name_list)){
  age_comps_real <-list()
  n_nom_real_trt <-list()
  n_nom_sim_trt <-list()
  yrs <-sort(unique(dplyr::filter(age_data, species_description==sp_name_list[sp])$return_yr))

  for(s in 1:length(samp_rate)){
    age_comps_b <-list()
    n_nom_sim_b <-list()
    n_nom_real_b <-list()
    print(samp_rate[[s]])
    
    ####################################################################
    ### Calc real age comps & sim age comps boots for all years combined
    age_data_sp <-dplyr::filter(age_data, species_description==sp_name_list[sp])
    age_data_sp_tab_real <-table(age_data_sp$gilbert_rich_int)

    age_vec_sp   <-c(as.numeric(ifelse(is.na(age_data_sp_tab_real["1"])==FALSE,age_data_sp_tab_real["1"],0)),
                     as.numeric(ifelse(is.na(age_data_sp_tab_real["2"])==FALSE,age_data_sp_tab_real["2"],0)),
                     as.numeric(ifelse(is.na(age_data_sp_tab_real["3"])==FALSE,age_data_sp_tab_real["3"],0)),
                     as.numeric(ifelse(is.na(age_data_sp_tab_real["4"])==FALSE,age_data_sp_tab_real["4"],0)),
                     as.numeric(ifelse(is.na(age_data_sp_tab_real["5"])==FALSE,age_data_sp_tab_real["5"],0)),
                     as.numeric(ifelse(is.na(age_data_sp_tab_real["6"])==FALSE,age_data_sp_tab_real["6"],0)))
    
    age_comps_real_sp[[y]] <-age_vec_sp/sum(age_vec_sp)
    
    samp_dates_sp <-unique(age_data$month_date)
    age_data_sim_sp <-list()

    
    for(b in 1:n_boot){
      print(b)
      set.seed(bootest.seed[b])
      age_comps_sim <-list()
      n_nom_y <-list()
      
      
      for(sd in 1:length(samp_dates_sp)){
        age_data_sp_sd <-dplyr::filter(age_data, month_date==samp_dates[sd])
        age_data_sim_sp[[sd]] <-sample(age_data_sp_sd$gilbert_rich_int,size=floor(round(nrow(age_data_sp_sd)*samp_rate[s])),replace=TRUE)
      }
      
      names(age_data_sim_sp) <-samp_dates_sp
      age_data_sim_sp_df <-map_df(age_data_sim_sp,~as.data.frame(.))
      age_data_tab_sim_sp <-table(age_data_sim_sp_df)
      
      age_vec_sim_sp<-c(as.numeric(ifelse(is.na(age_data_tab_sim_sp["1"])==FALSE,age_data_tab_sim_sp["1"],0)),
                        as.numeric(ifelse(is.na(age_data_tab_sim_sp["2"])==FALSE,age_data_tab_sim_sp["2"],0)),
                        as.numeric(ifelse(is.na(age_data_tab_sim_sp["3"])==FALSE,age_data_tab_sim_sp["3"],0)),
                        as.numeric(ifelse(is.na(age_data_tab_sim_sp["4"])==FALSE,age_data_tab_sim_sp["4"],0)),
                        as.numeric(ifelse(is.na(age_data_tab_sim_sp["5"])==FALSE,age_data_tab_sim_sp["5"],0)),
                        as.numeric(ifelse(is.na(age_data_tab_sim_sp["6"])==FALSE,age_data_tab_sim_sp["6"],0)))

      
      n_nom_real_b[[b]] <-sum(age_vec_sim_sp)
      age_comps_sim_b[[b]] <-age_vec_sim_sp/sum(age_vec_sim_sp)
      
      ###########################################################
      ### Calc real age comps & sim age comps boots for each year
      for(y in 1:length(yrs)){
        ### Calc real age comps for each year
        age_data_sp_y <-dplyr::filter(age_data, species_description==sp_name_list[sp] & return_yr==yrs[y])
        age_data_tab_real <-table(age_data_sp_y$gilbert_rich_int)
    
        age_vec   <-c(as.numeric(ifelse(is.na(age_data_tab_real["1"])==FALSE,age_data_tab_real["1"],0)),
                        as.numeric(ifelse(is.na(age_data_tab_real["2"])==FALSE,age_data_tab_real["2"],0)),
                        as.numeric(ifelse(is.na(age_data_tab_real["3"])==FALSE,age_data_tab_real["3"],0)),
                        as.numeric(ifelse(is.na(age_data_tab_real["4"])==FALSE,age_data_tab_real["4"],0)),
                        as.numeric(ifelse(is.na(age_data_tab_real["5"])==FALSE,age_data_tab_real["5"],0)),
                        as.numeric(ifelse(is.na(age_data_tab_real["6"])==FALSE,age_data_tab_real["6"],0)))
        
        age_comps_real[[y]] <-age_vec/sum(age_vec)
        
        samp_dates <-unique(age_data_sp_y$month_date)
        age_data_sim <-list()
        
        for(sd in 1:length(samp_dates)){
          age_data_sp_y_sd <-dplyr::filter(age_data_sp_y, month_date==samp_dates[sd])
          age_data_sim[[sd]] <-sample(age_data_sp_y_sd$gilbert_rich_int,size=floor(round(nrow(age_data_sp_y_sd)*samp_rate[s])),replace=TRUE)
        }
        
        names(age_data_sim) <-samp_dates
        age_data_sim_df <-map_df(age_data_sim,~as.data.frame(.))
        age_data_tab_sim <-table(age_data_sim_df)
        
        age_vec_sim<-c(as.numeric(ifelse(is.na(age_data_tab_sim["1"])==FALSE,age_data_tab_sim["1"],0)),
                       as.numeric(ifelse(is.na(age_data_tab_sim["2"])==FALSE,age_data_tab_sim["2"],0)),
                       as.numeric(ifelse(is.na(age_data_tab_sim["3"])==FALSE,age_data_tab_sim["3"],0)),
                       as.numeric(ifelse(is.na(age_data_tab_sim["4"])==FALSE,age_data_tab_sim["4"],0)),
                       as.numeric(ifelse(is.na(age_data_tab_sim["5"])==FALSE,age_data_tab_sim["5"],0)),
                       as.numeric(ifelse(is.na(age_data_tab_sim["6"])==FALSE,age_data_tab_sim["6"],0)))
  
        n_nom_y[[y]] <-sum(age_vec_sim)
        age_comps_sim[[y]] <-age_vec_sim/sum(age_vec_sim)
      }
      names(n_nom_y) <-yrs
      n_nom_sim_b[[b]] <-n_nom_y
      names(age_comps_real) <-yrs
      names(age_comps_sim) <-yrs
      age_comps_b[[b]] <-age_comps_sim
    }
    n_nom_real_trt[[s]] <-n_nom_real_b
    n_nom_sim_trt[[s]] <-n_nom_sim_b
    age_comps_sim_trt[[s]] <-age_comps_b
  }
  names(n_nom_real_trt) <-samp_rate
  names(n_nom_sim_trt) <-samp_rate
  n_nom_sp[[sp]] <-n_nom_trt
  age_comps_real_sp[[sp]] <-age_comps_real
  names(age_comps_sim_trt) <-samp_rate
  age_comps_sim_sp[[sp]] <-age_comps_sim_trt
}
names(n_nom_real_sp) <-sp_name_list
names(n_nom_sim_sp) <-sp_name_list
names(age_comps_real_sp) <-sp_name_list
names(age_comps_sim_sp) <-sp_name_list

for(sp in 1:length(sp_name_list)){
  for(s in 1:length(samp_rate)){
    if(s==1){age_comps_sim_df <-cbind(as.data.frame(t(map_df(age_comps_sim_sp[[sp]][[s]],~as.data.frame(.)))),yrs)}
    if(s>1){
      temp <-cbind(as.data.frame(t(map_df(age_comps_sim_sp[[sp]][[s]],~as.data.frame(.)))),yrs)
      age_comps_sim_df <-rbind(temp, age_comps_sim_df)
    }
  }
}

# Table of samples taken per sampling date
sample_tab <-t(table(age_data[,c("return_yr","month_date")]))
tab_with_sums <- addmargins(sample_tab, 1) # 1 specifies adding row margins (sums of columns)
tab_with_sums %>% 
  kbl(row.names = TRUE, caption = "Table 1. Count of samples taken on each sampling date.") %>%
  kable_classic(full_width = F, position = "left", html_font = "Cambria", fixed_thead = T)

# Plot of sample count by sampling date
ggplot(data = age_data[,c("return_yr","month_date")], aes(x = month_date)) +
  geom_bar(fill = "steelblue", color = "white")+
  facet_wrap(return_yr~.)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Angle x-axis labels for readability

age_comps_real <-rbindlist(age_comps_real_sp, fill=FALSE)
age_comps_real %>% 
  kbl(row.names = TRUE, caption = "Table 2. Real age compositions calcualted for each year of sampling.") %>%
  kable_classic(full_width = F, position = "left", html_font = "Cambria", fixed_thead = T)

# print(Sys.time())


```

## Calculate data weights (n_eff) & plot

```{r inputSS}
# Standardizes Year x Age matrix to proportional values across age classes w/i each year
# standardize = function(mat) t(mat) / rep(1,ncol(mat))%o%rowSums(mat)

# Calculate N_bt
calculate_n = function(boot_ct, obs_ct){
  # boot_ct  = standardize(boot_ct)
  # obs_ct   = standardize(obs_ct)
  num_ct   = boot_ct*(1-boot_ct)
  denom_ct = (boot_ct - obs_ct)^2
  colSums(num_ct)/colSums(denom_ct)
}

# Calculate harmonic mean of N_bt
harmonic.mean = function(vec) 1/mean(1/vec)

# Set 'observed' proportions to real annual age comps, calc N_bt
# Could use mean of specimen bootstrap replicates if implemented 
samp_suite_sp <-list()

for(sp in 1:length(sp_name_list)){
  samp_suite <-list()
  
  for(trt in 1:length(samp_rate)){
    N_bt = matrix(NA, nrow=n_boot, ncol=length(yrs))
    colnames(N_bt) <-yrs
    print(samp_rate[[trt]])
    
    for(y in 1:length(yrs)){
      for(bI in 1:n_boot){
        N_bt[bI,y] = t(calculate_n(boot_ct=map_df(age_comps_sim_sp[[sp]][[trt]][[bI]][[y]],~as.data.frame(.)), 
                                  obs_ct=map_df(age_comps_real_sp[[sp]][[y]],~as.data.frame(.))))
      }
    }
    # # Calculate input sample size (harmonic mean across bootstrap samples)
    Nsamp_t = array(NA,dim=length(yrs))
    Nsamp_t = apply(N_bt, MARGIN=2, FUN=harmonic.mean)
    samp_suite[[trt]] <-Nsamp_t

  }
  names(samp_suite) <-samp_rate
  samp_suite_sp[[sp]] <-samp_suite
}
names(samp_suite_sp) <-sp_name_list

fig_dat                <-map_df(samp_suite_sp,~as.data.frame(.))
colnames(fig_dat)      <-samp_rate
fig_dat$return_yr      <-rownames(fig_dat)
fig_dat_long           <-melt(setDT(fig_dat), id.vars = c("return_yr"), variable.name = "trt")
colnames(fig_dat_long) <-c("return_yr", "trt", "n_eff")
fig_dat_long           <-fig_dat_long[order(fig_dat_long$return_yr,fig_dat_long$trt),]

for(sp in 1:length(sp_name_list)){
  for(s in 1:length(samp_rate)){
    if(s==1){
      n_nom_df <-rbindlist(n_nom_sp[[sp]][[s]], fill=FALSE)
      n_nom_df$trt <-rep(samp_rate[[s]], n_boot)
    }
    if(s>1){
      temp <-rbindlist(n_nom_sp[[sp]][[s]], fill=FALSE)
      temp$trt <-rep(samp_rate[[s]], n_boot)
      n_nom_df <-rbind(n_nom_df, temp)
    }
  }
}

n_nom_df_long           <-melt(setDT(n_nom_df), id.vars = c("trt"), variable.name = "return_yr")
colnames(n_nom_df_long) <-c("trt","return_yr","n_nom")
n_nom_means             <-aggregate(n_nom ~ trt + return_yr, data = n_nom_df_long, FUN = mean)
n_nom_means$return_yr   <-as.character(n_nom_means$return_yr)
n_nom_means$trt         <-as.factor(n_nom_means$trt)
fig_dat_long            <-left_join(fig_dat_long, n_nom_means, by = c("return_yr","trt"))
fig_dat_long$n_diff     <-fig_dat_long$n_nom - fig_dat_long$n_eff

##### n_eff ~ n_nom
# Plot n_nom_means & n_eff across all return years
ggplot(data=fig_dat_long, aes(x=n_nom,y=n_eff,group=return_yr,fill=return_yr))+
  geom_point(size=3,pch=21)+
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed")

# Plot n_nom_means & n_eff for each return year
ggplot(data=fig_dat_long, aes(x=n_nom,y=n_eff,group=return_yr))+
  geom_point()+
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed")+
  facet_wrap(return_yr~.)

##### n_diffs
# Plot difference between n_nom & n_eff across trt (all yrs combined)
plot(x=fig_dat_long$trt,y=fig_dat_long$n_diff, main="n_diff = n_samp - n_eff")

# Plot difference between n_nom & n_eff across trt (yr facets)
ggplot(data=fig_dat_long, aes(x=trt,y=n_diff,group=return_yr))+
  geom_point()+
  geom_abline(intercept = 0, slope = 0, color = "red", linetype = "dashed")+
  facet_wrap(return_yr~.)

# Plot difference between n_nom & n_eff across n_nom
ggplot(data=fig_dat_long, aes(x=n_nom,y=n_diff,group=return_yr))+
  geom_point()+
  geom_abline(intercept = 0, slope = 0, color = "red", linetype = "dashed")+
  facet_wrap(return_yr~.)

```

## Log-log GLMM fit of n_eff \~ n_samp

```{r glmm}
mod <-glmmTMB( log(n_eff) ~ 1 + log(n_samp)*( 1 + log(n_samp)),data=fig_dat)
print(mod)
LOGLOGslope <-mod[["fit"]][["par"]][[2]]
mod_pred <-ggpredict(mod, terms = "n_samp")
colnames(mod_pred) <-c("n_samp", "n_eff", "std.error", "conf.low", "conf.high", "group")

base_plot <- ggplot(data = fig_dat, aes(x = n_samp, y = n_eff)) +
      geom_point()

base_plot +
  geom_line(data = mod_pred, aes(x = n_samp, y = n_eff), color = "blue", size = 1) +
  geom_ribbon(data = mod_pred, aes(x = n_samp, ymin = conf.low, ymax = conf.high), alpha = 0.2, fill = "lightblue")



```
